---
title: "Indicator catalogue"
---

## Accessed datasets

This page summarizes the supported catalogue of indicators. The current version
accesses datasets from:

### Monthly Resolution

**1. Copernicus Climate Change Service (C3S)**:  

- [ERA5-Land Monthly Averaged Data (1950–Present)](https://cds.climate.copernicus.eu/datasets/reanalysis-era5-land-monthly-means?tab=overview)
- [ERA5 Monthly Averaged Data on Single Levels (1940–Present)](https://cds.climate.copernicus.eu/datasets/reanalysis-era5-single-levels-monthly-means?tab=overview)

### Daily Resolution

- [ERA5-Land post-processed daily statistics from 1950 to present](https://cds.climate.copernicus.eu/datasets/derived-era5-land-daily-statistics?tab=overview)
- [ERA5 post-processed daily statistics on single levels from 1940 to present](https://cds.climate.copernicus.eu/datasets/derived-era5-single-levels-daily-statistics?tab=overview)

### Future datasets include:

- [Thermal comfort indices derived from ERA5 reanalysis](https://cds.climate.copernicus.eu/datasets/derived-utci-historical?tab=overview)
- [ERA5 hourly data on single levels from 1940 to present](https://cds.climate.copernicus.eu/datasets/reanalysis-era5-single-levels?tab=overview)
- [ERA5-Land hourly data from 1950 to present](https://cds.climate.copernicus.eu/datasets/reanalysis-era5-land?tab=overview)
- [Fire burned area from 2001 to present derived from satellite observations](https://cds.climate.copernicus.eu/datasets/satellite-fire-burned-area?tab=overview)
- [Emission Database for Global Atmospheric Research (EDGAR)](https://edgar.jrc.ec.europa.eu/)
- [CAMS global greenhouse gas reanalysis (EGG4)](https://ads.atmosphere.copernicus.eu/datasets/cams-global-ghg-reanalysis-egg4?tab=overview)

## Dataset Attributes

```{r ERA5-monthly, echo=FALSE, warning=FALSE}
library(knitr)

# Create dataset attributes table for monthly resolution
dataset_attributes <- data.frame(
  Attribute = c("Projection", "Spatial Resolution", "Spatial Coverage (Extent)", "Temporal Resolution", "Temporal Coverage", "Update Frequency"),
  `ERA5-Land Monthly Averaged Data` = c("WGS84 (EPSG:4326)", "~9 km (~0.1°)", "Global (-90° to 90°, -180° to 180°)", "Monthly", "1950 to present", "Monthly"),
  `ERA5 Monthly Averaged Data on Single Levels` = c("WGS84 (EPSG:4326)", "~31 km (~0.25°)", "Global (-90° to 90°, -180° to 180°)", "Monthly", "1940 to present", "Monthly")
)

# Render as table
kable(dataset_attributes, format = "markdown")
```

```{r ERA5-daily, echo=FALSE, warning=FALSE}
library(knitr)

# Create dataset attributes table for daily resolution
dataset_attributes_daily <- data.frame(
  Attribute = c("Projection", "Spatial Resolution", "Spatial Coverage (Extent)", "Temporal Resolution", "Temporal Coverage", "Update Frequency"),
  `Derived ERA5 Land Daily Statistics` = c("WGS84 (EPSG:4326)", "~9 km (~0.1°)", "Global (-90° to 90°, -180° to 180°)", "Daily", "As available", "Daily"),
  `Derived ERA5 Single Levels Daily Statistics` = c("WGS84 (EPSG:4326)", "~31 km (~0.25°)", "Global (-90° to 90°, -180° to 180°)", "Daily", "As available", "Daily")
)

kable(dataset_attributes_daily, format = "markdown")
```

## Supported indicators

### Monthly resolution indicators

```{r echo=FALSE, warning=FALSE}
# Create variables table
variables_table <- data.frame(
  Variable = c("2m_temperature", "total_precipitation", "10m_u_component_of_wind", "10m_v_component_of_wind",
               "10m_wind_speed", "total_cloud_cover", "leaf_area_index_high_vegetation", "leaf_area_index_low_vegetation", "snowfall"),
  Description = c("Air temperature at 2 meters above the surface (°C)",
                  "Total accumulated precipitation (m)",
                  "Eastward component of wind speed at 10 meters (m/s)",
                  "Northward component of wind speed at 10 meters (m/s)",
                  "Wind speed magnitude at 10 meters (m/s)",
                  "Fraction of sky covered by clouds (0-1)",
                  "Leaf area index for high vegetation (m²/m²)",
                  "Leaf area index for low vegetation (m²/m²)",
                  "Total accumulated snowfall (m)"),
  `ERA5-Land` = c("✔", "✔", "✔", "✔", "", "", "✔", "✔", "✔"),
  `ERA5 Single Levels` = c("✔", "", "✔", "✔", "✔", "✔", "✔", "✔", "")
)

# Render as table
kable(variables_table, format = "markdown")

```

### Daily resolution indicators

```{r echo=FALSE, warning=FALSE}
# Create variables table for daily resolution
daily_variables_table <- data.frame(
  Variable = c("2m_temperature"),
  Description = c("Air temperature at 2 m above the surface (°C)"),
  `Derived ERA5 Land Daily Statistics` = c("✔"),
  `Derived ERA5 Single Levels Daily Statistics` = c("✔")
)

kable(daily_variables_table, format = "markdown")

```