---
title: "Performance"
---

## How long will the package run on an average laptop?

Here, you can explore performance of the `gxc`-package in terms of execution time. We have tested several specifications and varied the most important parameters which might affect the processing time and which are representative of typical specifications in previous studies: 

1. **Sample size**: 10-5000 
2. **Spatial extent**: City, national, continental, worldwide 
3. **Focal period**: Averages for month, season, and year
4. **Baseline period**: No baseline, 10y baseline, 30y baseline.

We are accessing the `2m_temperature` indicator from the `ERA5-Land` dataset for this exercise. The code was run on an average office laptop (Intel Core i7-10510U CPU, 16GB RAM, Windows 10). Below, you find the detailed code for the sample generation and execution.

```{r packages, include = FALSE}
# Load packages
library(sf)
library(tidyverse)
library(plotly)
library(rnaturalearth)
library(gridExtra)
library(keyring)
library(tictoc)
library(gxc)

```


```{r create-grid, echo = FALSE}
# Load shapefiles and prepare grids ---------------------------------------

# Retrieve Germany's bounding box
germany <- ne_countries(country = "Germany", 
                        scale = "medium", 
                        returnclass = "sf") 
  
# Generate the grid
grid_national <- sf::st_make_grid(
  st_as_sfc(st_bbox(germany)), 
  n = c(100,100)
  ) 
  
grid_national <- st_sf(geometry = grid_national)

# Assign WGS84 CRS
st_crs(grid_national) <- 4326

# Define the four corner cells (by relying on the indexing of the grid cell)
ll_index <- 1
lr_index <- 100
ul_index <- 9901
ur_index <- 10000

corner_indices <- c(ll_index, lr_index, ul_index, ur_index)


# Generate sample grids ---------------------------------------------------

# Sample-function (always including the four corner cells).
sample_grid <- function(n, my_grid, extent_polygons) {
  if(n < 4) stop("n must be at least 4 to include all corner cells.")
  
  # Calculate n-corner polygons
  n_random <- n - length(extent_polygons)
  
  # Exclude the corner cells from the random sample
  remaining <- my_grid[-extent_polygons, ]
  
  # Randomly sample from the remaining cells
  random_sample <- remaining[sample(nrow(remaining), n_random), ]
  
  # Combine the corner cells with the random sample
  sampled <- rbind(my_grid[extent_polygons, ], random_sample)
  return(sampled)
}

# Sample grids with various n
sample_sizes <- c(10, 20, 50, 100, 200, 500, 1000, 2000, 5000)

samples_list <- list()

# Loop over the sample sizes
for (i in seq_along(sample_sizes)) {
  n <- sample_sizes[i]

  samples_list[[i]] <- sample_grid(n, grid_national, corner_indices)
}

names(samples_list) <- paste0("sample_", sample_sizes)

```


```{r grid-plot, include = FALSE}

# Visualize
plots <- lapply(seq_along(sample_sizes), function(i) {
  n <- sample_sizes[i]
  sampled_sf <- samples_list[[i]]
  
  ggplot() +
    geom_sf(data = sampled_sf, 
            fill = "#c994c7", 
            color = "#dd1c77", 
            size = 0.3) +
    geom_sf(data = germany, fill = NA, color = "#2c7fb8", size = 0.8) +
    ggtitle(paste("Sample size:", n)) +
    theme_minimal()
})

gridExtra::grid.arrange(grobs = plots[-c(1:3)], ncol = 3)

```

```{r execute-time, include = FALSE}
# # Loop over poly_link-function and measure elapsed time -------------------
# 
# # Add random time identifier in grids
# samples_list <- lapply(samples_list, function(grid) {
#   grid$date_raw <- "12-2020"
#   grid
# })
# 
# # Set up API access
# api_key <- Sys.getenv("WF_API_KEY")
# 
# keyring::key_set_with_value(service = "wf_api_key", password = api_key)
# 
# # Set up new dataframe measuring elapsed time
# results_df <- data.frame(
#   spatial_extent = character(),
#   sample_size = integer(),
#   focal_period = character(),
#   seconds = numeric(),
#   stringsAsFactors = FALSE
# )
# 
# # Loop over the list of samples using tictoc to measure execution time.
# time_span_values <- c(0,2,11)
# 
# for (i in seq_along(samples_list)) {
#   sample_size <- sample_sizes[i]
#   sample_polygons <- samples_list[[i]]
#   
#   for (ts in time_span_values) {
#   
#   # Start timing using tic().
#   tic()  
#   # Run poly_link function
#   poly_result <- poly_link(
#     indicator = "2m_temperature",
#     data = sample_polygons,
#     date_var = "date_raw",
#     time_span = ts,
#     time_lag = 0,
#     baseline = FALSE,
#     order = "my",
#     path = "./data",
#     catalogue = "reanalysis-era5-land-monthly-means",
#     by_hour = FALSE,
#     keep_raw = FALSE)
#   # Stop timing and capture the elapsed time.
#   toc_out <- toc(quiet = TRUE)
#   
#   # The output toc_out is a list with the start and stop times. Compute the elapsed time.
#   elapsed_time <- toc_out$toc - toc_out$tic
#   
#   focal_period <- if (ts == 0) {
#     "Month"
#   } else if (ts == 2) {
#     "Season"
#   } else if (ts == 11) {
#     "Year"
#   } else {
#     NA
#   }
#   
#   # Append the results to the results_df data frame.
#   results_df <- rbind(results_df, data.frame(
#     spatial_extent = "Germany",
#     sample_size = sample_size,
#     focal_period = focal_period,
#     seconds = elapsed_time,
#     stringsAsFactors = FALSE
#   ))
#   
#   # Save as rds
#   saveRDS(results_df, file = "./data/performance_results.rds")
#   }
# }

```

```{r time-plot, echo = FALSE, message = FALSE}
results_df <- readRDS("./data/performance_results.rds")

# split data based on groups for tabs
df_list <- results_df %>% 
  group_by(spatial_extent) %>% 
  group_split()

# Default ggplot
p <- ggplot() +
  labs(
    title = "Execution Time vs Sample Size",
    x = "Sample Size",
    y = "Execution Time (minutes)",
    color = "Focal Period"
  ) +
  scale_color_manual(values = c("month" = "#dd1c77",
                                "seasonal" = "#225ea8",
                                "yearly" = "#7fcdbb"))+
  theme_minimal()

# Looping through list of subsets
for (d in df_list) {
  p <- p +
    geom_point(data = d, aes(x = sample_size, y = seconds / 60, color = focal_period)) +
    geom_smooth(data = d, aes(x = sample_size, y = seconds / 60, color = focal_period),
                method = "loess", se = FALSE)
}

# Convert to plotly object
p_plotly <- ggplotly(p)

# Update layout

# Create buttons
# Spatial extent
spatial_extent_buttons <- list(
  list(
    method = "restyle",
    args = list("visible", list(TRUE, TRUE)),
    label = "Germany"
  ),
  list(
    method = "restyle",
    args = list("visible", list(FALSE, FALSE)),
    label = "Europe"
  )
)

# Baseline period
baseline_buttons <- list(
  list(
    method = "restyle",
    args = list("line.dash", "solid"),
    label = "No baseline"
  ),
  list(
    method = "restyle",
    args = list("line.dash", "dot"),
    label = "10y baseline"
  ),
  list(
    method = "restyle",
    args = list("line.dash", "dash"),
    label = "30y baseline"
  )
)

# Adjust layout accordingly
p_plotly <- p_plotly %>% layout(
  title = list(
    text = "Execution Time vs Sample Size",
    y = 0.85
  ),
  # Increase the top margin to give more space
  margin = list(t = 150),
  updatemenus = list(
    list(
      type = "dropdown",
      active = 0,
      buttons = spatial_extent_buttons,
      x = 0, y = 1.12,
      xanchor = "left",
      yanchor = "top"
    ),
    list(
      type = "dropdown",
      active = 0,
      buttons = baseline_buttons,
      x = 0.2, y = 1.12,
      xanchor = "left",
      yanchor = "top"
    )
  ),
  annotations = list(
    list(
      x = 0, y = 1.22,
      xref = "paper", yref = "paper",
      text = "Spatial extent",
      showarrow = FALSE,
      font = list(size = 14)
    ),
    list(
      x = 0.2, y = 1.22,
      xref = "paper", yref = "paper",
      text = "Baseline period",
      showarrow = FALSE,
      font = list(size = 14)
    )
  )
)

# Display the interactive plot.
p_plotly

```


## Replication code

```{r packages2, warning = FALSE, message = FALSE}
# Load packages
library(sf)
library(tidyverse)
library(plotly)
library(rnaturalearth)
library(gridExtra)
library(keyring)
library(tictoc)
library(gxc)

```

## Generate random sample

```{r create-grid2}
# Load shapefiles and prepare grids ---------------------------------------

# Retrieve Germany's bounding box
germany <- ne_countries(country = "Germany", 
                        scale = "medium", 
                        returnclass = "sf") 
  
# Generate the grid
grid_national <- sf::st_make_grid(
  st_as_sfc(st_bbox(germany)), 
  n = c(100,100)
  ) 
  
grid_national <- st_sf(geometry = grid_national)

# Assign WGS84 CRS
st_crs(grid_national) <- 4326

# Define the four corner cells (by relying on the indexing of the grid cell)
ll_index <- 1
lr_index <- 100
ul_index <- 9901
ur_index <- 10000

corner_indices <- c(ll_index, lr_index, ul_index, ur_index)


# Generate sample grids ---------------------------------------------------

# Sample-function (always including the four corner cells).
sample_grid <- function(n, my_grid, extent_polygons) {
  if(n < 4) stop("n must be at least 4 to include all corner cells.")
  
  # Calculate n-corner polygons
  n_random <- n - length(extent_polygons)
  
  # Exclude the corner cells from the random sample
  remaining <- my_grid[-extent_polygons, ]
  
  # Randomly sample from the remaining cells
  random_sample <- remaining[sample(nrow(remaining), n_random), ]
  
  # Combine the corner cells with the random sample
  sampled <- rbind(my_grid[extent_polygons, ], random_sample)
  return(sampled)
}

# Sample grids with various n
sample_sizes <- c(10, 20, 50, 100, 200, 500, 1000, 2000, 5000)

samples_list <- list()

# Loop over the sample sizes
for (i in seq_along(sample_sizes)) {
  n <- sample_sizes[i]

  samples_list[[i]] <- sample_grid(n, grid_national, corner_indices)
}

names(samples_list) <- paste0("sample_", sample_sizes)

```


```{r grid-plot2}

# Visualize
plots <- lapply(seq_along(sample_sizes), function(i) {
  n <- sample_sizes[i]
  sampled_sf <- samples_list[[i]]
  
  ggplot() +
    geom_sf(data = sampled_sf, 
            fill = "#c994c7", 
            color = "#dd1c77", 
            size = 0.3) +
    geom_sf(data = germany, fill = NA, color = "#2c7fb8", size = 0.8) +
    ggtitle(paste("Sample size:", n)) +
    theme_minimal()
})

gridExtra::grid.arrange(grobs = plots[-c(1:3)], ncol = 3)

```

## Calculate execution time for each specification

```{r execute-time2}
# # Loop over poly_link-function and measure elapsed time -------------------
# 
# # Add random time identifier in grids
# samples_list <- lapply(samples_list, function(grid) {
#   grid$date_raw <- "12-2020"
#   grid
# })
# 
# # Set up API access
# api_key <- Sys.getenv("WF_API_KEY")
# 
# keyring::key_set_with_value(service = "wf_api_key", password = api_key)
# 
# # Set up new dataframe measuring elapsed time
# results_df <- data.frame(
#   spatial_extent = character(),
#   sample_size = integer(),
#   focal_period = character(),
#   seconds = numeric(),
#   stringsAsFactors = FALSE
# )
# 
# # Loop over the list of samples using tictoc to measure execution time.
# time_span_values <- c(0,2,11)
# 
# for (i in seq_along(samples_list)) {
#   sample_size <- sample_sizes[i]
#   sample_polygons <- samples_list[[i]]
#   
#   for (ts in time_span_values) {
#   
#   # Start timing using tic().
#   tic()  
#   # Run poly_link function
#   poly_result <- poly_link(
#     indicator = "2m_temperature",
#     data = sample_polygons,
#     date_var = "date_raw",
#     time_span = ts,
#     time_lag = 0,
#     baseline = FALSE,
#     order = "my",
#     path = "./data",
#     catalogue = "reanalysis-era5-land-monthly-means",
#     by_hour = FALSE,
#     keep_raw = FALSE)
#   # Stop timing and capture the elapsed time.
#   toc_out <- toc(quiet = TRUE)
#   
#   # The output toc_out is a list with the start and stop times. Compute the elapsed time.
#   elapsed_time <- toc_out$toc - toc_out$tic
#   
#   focal_period <- if (ts == 0) {
#     "Month"
#   } else if (ts == 2) {
#     "Season"
#   } else if (ts == 11) {
#     "Year"
#   } else {
#     NA
#   }
#   
#   # Append the results to the results_df data frame.
#   results_df <- rbind(results_df, data.frame(
#     spatial_extent = "Germany",
#     sample_size = sample_size,
#     focal_period = focal_period,
#     seconds = elapsed_time,
#     stringsAsFactors = FALSE
#   ))
#   
#   # Save as rds
#   saveRDS(results_df, file = "./data/performance_results.rds")
#   }
# }

```

## Visualize

```{r time-plot2}
results_df <- readRDS("./data/performance_results.rds")

# split data based on groups for tabs
df_list <- results_df %>% 
  group_by(spatial_extent) %>% 
  group_split()

# Default ggplot
p <- ggplot() +
  labs(
    title = "Execution Time vs Sample Size",
    x = "Sample Size",
    y = "Execution Time (minutes)",
    color = "Focal Period"
  ) +
  scale_color_manual(values = c("month" = "#dd1c77",
                                "seasonal" = "#225ea8",
                                "yearly" = "#7fcdbb"))+
  theme_minimal()

# Looping through list of subsets
for (d in df_list) {
  p <- p +
    geom_point(data = d, aes(x = sample_size, y = seconds / 60, color = focal_period)) +
    geom_smooth(data = d, aes(x = sample_size, y = seconds / 60, color = focal_period),
                method = "loess", se = FALSE)
}

# Convert to plotly object
p_plotly <- ggplotly(p)

# Update layout

# Create buttons
# Spatial extent
spatial_extent_buttons <- list(
  list(
    method = "restyle",
    args = list("visible", list(TRUE, TRUE)),
    label = "Germany"
  ),
  list(
    method = "restyle",
    args = list("visible", list(FALSE, FALSE)),
    label = "Europe"
  )
)

# Baseline period
baseline_buttons <- list(
  list(
    method = "restyle",
    args = list("line.dash", "solid"),
    label = "No baseline"
  ),
  list(
    method = "restyle",
    args = list("line.dash", "dot"),
    label = "10y baseline"
  ),
  list(
    method = "restyle",
    args = list("line.dash", "dash"),
    label = "30y baseline"
  )
)

# Adjust layout accordingly
p_plotly <- p_plotly %>% layout(
  title = list(
    text = "Execution Time vs Sample Size",
    y = 0.85
  ),
  # Increase the top margin to give more space
  margin = list(t = 150),
  updatemenus = list(
    list(
      type = "dropdown",
      active = 0,
      buttons = spatial_extent_buttons,
      x = 0, y = 1.12,
      xanchor = "left",
      yanchor = "top"
    ),
    list(
      type = "dropdown",
      active = 0,
      buttons = baseline_buttons,
      x = 0.2, y = 1.12,
      xanchor = "left",
      yanchor = "top"
    )
  ),
  annotations = list(
    list(
      x = 0, y = 1.22,
      xref = "paper", yref = "paper",
      text = "Spatial extent",
      showarrow = FALSE,
      font = list(size = 14)
    ),
    list(
      x = 0.2, y = 1.22,
      xref = "paper", yref = "paper",
      text = "Baseline period",
      showarrow = FALSE,
      font = list(size = 14)
    )
  )
)

# Display the interactive plot.
# p_plotly

```